\chapter{Background and Related Work}

The creation of a metric builder page and an API for the MATTERS site is a result of two prior project groups and additional efforts by WPI students who worked to establish the initial features and the design framework for MATTERS to exist. The initial development of MATTERS came from a team of students completing their Interactive Qualifying Project (IQP) with the aid of other WPI graduate student teams. This team of IQP students researched and came to the resulting decisions regarding the front-end visualization for the MATTERS dashboard [1]. This includes the inclusion of decisions regarding colors, and visualization types such as charts and tables, that would be used to the represent the data within the system in a manner that was both intuitive for the users and a quick way to show analytics. In parallel, the graduate student teams worked to develop the back-end for MATTERS. They made it possible to extract the data wanted from various sites and then parse and clean this retrieved data, before then uploading it to a database. This data would be called upon to be shown in the various visual methods for the users [2].

Following this initial front-end and back-end design and implementation of the MATTERS dashboard, a team of students working on their Major Qualifying Project (MQP) for WPI then worked to improve the administration center and the data integration for the dashboard. This involved the data integration pipeline manager as well as the Administration Center’s development to allow for MHTC and other administrators who had access to be able to easily upload their desired data and metrics, as well as an easy way to view this data on the administrative end. This MQP team was able to integrate 20 new data sources for the MATTERS dashboard, make the Administration Center useable for non-technical users, and improve upon the look and feel of the MATTERS dashboard as a whole [2].

\section{System Architecture}

\section{Requirements}

The main goal of creating an API for MATTERS as well as creating the Metric Buidler page for the site and what it would entail, required looking into other similar features of different sites and making decisions regarding implementation, design and security features for these aspects of the MATTERS site.

\subsection{API and Documentation}

An API is an Application Programming Interface. APIs are tools that allow other programs to interact with the software program that the API belongs to without having to give someone complete access to your codebase. In a sense they expose some of the internal workings of software to the public but in a way that is limited by what the API chooses to provide. APIs aid in providing a way to share a potentially large amount of data in an efficient way. According to the chief data officer of Philadelphia, Mark Headd, APIs “allow a specific audience to use data more quickly, easily, and efficiently when they are looking to do something specific with the information.” APIs are beneficial for not only keeping certain aspects of code private, but also for saving users time. Even in cases of open source programs where all of the code for a program is visible, the number of code can be so large that it is inefficient to search through the whole codebase for specific data [3]. 

APIs will be beneficial from both a business aspect and for programs that aim share information and research with many people. An advantage of an API being an automated tool is that it can process a large number requests without any added cost and work from the developers. APIs also present the data requested in a manner that is useful and easily manipulated for the users’ purposes. For APIs whose goals are to share certain data across systems there is a great benefit efficiency wise, as many manual procedures are cut out by the APIs automatic generation. When these tasks are done manually, they waste time and are both laborious and repetitive, as well as being more prone to error; all of these things being costly on the developer end. While it is more time consuming and costly to develop an API initially, the overall benefit once it is developed is very clear [3].

Additionally, APIs not only encourage innovation and the ability to manipulate and extrapolate data by external collaborators, they provide a means to do so that is more secure for the group creating the API. They reduce the risk of how and what data is obtains and encourage good practice for how to properly manage the data available [4].

One important step to gaining the full benefits from an API is making sure that the API is secure. This can be achieved through requiring authentication and authorization of users before they gain access to the API’s features. It is good practice to ensure that the users are who they say they are and that they are given permission before using data or parts of a developer’s program. This also allows for a record of who exactly is accessing the data, in the event that there is an issue so that a specific user can be tracked down [5].

In addition to creating the API for a site, it is necessary to create good documentation to accompany the API, to make the API as user friendly and simplistic as possible. API documentation provides developers with everything that they need to know in a concise manner in order to use the API’s features with other applications. The API documentation needs to contain the classes, functions and other important aspects of the API [6]. The documentation guide allows developers to easily interact with a site’s API and their own code, as many developers style their code in their own unique way. 

\subsection{Metric Builder}

The main feature that is to be developed is the metric builder page which will be available for the registered users on the MATTERS website only. A user will be able to select any number of the metrics already available in the MATTERS database and combine their data to create their own personal formula or indicator. The user can then use their created formula to display the states rankings’ based on this indicator. This will allow users to see and evaluate states based on what they think is most important as a whole or combination of various factors, and interpret the data in their own way, simply. By allowing users to create their own data set from all the metrics available and then provide their own weights to each of these data values, we are allowing users to see the value that each state has based on how important they find a large number of metrics.

It is important to ensure that many different data types are able to be combined to create one final value. It is also important to consider the effects that certain data types will have on the final value received from the user’s indicator. For example, the MATTERS database contains rankings, percentages, nominal values, and other data types. Certain data types could end up dominating in a user created indicator when combined with significantly smaller value data types. There are also complexities involved with certain data types that are represented in the reverse of the majority of the data types. For example, both rankings and some other data types use numerical values in which a lower value indicates a better performance by the state, versus the majority of other metrics in which the larger the value is, the better a state is performing in that area. 

Finally, there are many metrics that do not contain a value for every single year that other metrics contain. Situations where one metric being used in the user’s metric builder equation does not contain values for years that another metric being used in the same equation need to be considered. There needs to be a uniform solution for what to do in these instances in a way that does not deceive the users of the indicator, as this would affect the calculations and following visualizations across years where these issues exist.

These factors need to be taken into consideration in order to provide accurate and understandable visualizations and results when users create their own indicator in the metric builder. It is for this reason good to consider special handling or removal of certain data types, as well as potential ways to normalize the final numerical values received from the formulas the users have created in the metric builder.

We looked at various economic indicators to come up with an idea for how to allow our users to combine sets of data to use for comparisons. One common approach in economic indicators is to use weighted averages [7]. For example, one of the most popular economic indicators, CPI, or the Consumer Price Index, uses weighted averages of data from hundreds of different consumer goods and services to measure inflation and deflation. Each of the goods considered by the CPI is weighted based on its importance [8].

Weighted averages work by taking each number in a given data set, and multiplying the value by its given weight. This will give you a new value reflecting the product of the data and its weight. Add all of these products together from the data set to get the total value. Then, add up the total of all the weights. Finally divide the total value by the total weight in order to receive the weighted average of your data set. The weights are used to show the importance that each specific item in a data set has in order to calculate an overall value of all the data together [9].
