
\chapter{Conclusion}

	In this project we have built two important services for the MATTERS site that enhances its capabilities and thus value. The API 
	provides a way for other computer-based systems to work with the MATTERS 
	system. Specifically, it is now possible to query MATTERS data directly via a software program 
	bypassing the presentation layer. This is extremely useful because it now empowers MATTERS' data warehouse to serve as a portal for important aggregated economic data.
	The other feature is the Metric Builder. 
	This service allows users to create their own personalized metrics based on 
	the existing metrics. This is useful as before users could only visualize the data of preexisting metrics, and MHTC and other users can now easily use their own personal idicators. 
	Users are able to work with their custom metrics seamlessly 
	in the Data Explorer as if they were just regular ones.  

	\section{Recommendations}
		
		Recommendations for other teams, either MQP or other teams, working on improving the MATTERS system in the future are as follows:
		
		\begin{itemize}
			\item MATTERS system
				\begin{enumerate}
					\item
						Let users define more complex equations for their user metrics.
						User should not be limited by simple weighted average formula.
					\item
						Let users suggest their own data or data sources. Right now 
						MATTERS' administrators and data management team are working on
						adding data. By involving users in this process MATTERS may have 
						more complete, accurate and up-to-date data.
					\item
						Role management. Right now the system supports regular users and API users. 
						Each user metric must have one and only one author. It would be useful 
						to merge these two user entities and introduce shared user metrics that are 
						not bound to specific users.
				\end{enumerate}
			\item Software development
				\begin{enumerate}
					\item
						Refactor the client side and server side code. Since the system 
						has been developed by a number of teams with different design 
						patterns and approaches, there is a significant technical debt. 
						It might be worth spending some time rewriting parts of the system 
						according to the latest coding and technical standards.
					\item
						Documentation. Right now there is a steep learning curve for new 
						developers who start working on the project. Good documentation would 
						dramatically decrease the time it takes for developers to start working 
						on the project.
				\end{enumerate}
		\end{itemize}
